SAT Competition 2020\tnoteref{title}

### 2021-06-18 paper/main.tex
working on section 6...

In the following, we provide an overview of the participating teams and solvers, 
and summarize new strategies which are implemented in the award-winning solvers of this competition. 
We start with a few remarks on the evolution of code-bases of well known SAT solvers in Section~\ref{sec:codebases}. 
Our review of the winning strategies in this competition starts with sequential solvers in the Main, Planning, and Incremental Library tracks in Section~\ref{sec:part:seq},
then we continue with winners of the Parallel track in Section~\ref{sec:part:par} and conclude with winning solvers in the Cloud track in Section~\ref{sec:part:cloud}.

\subsection{Code-Bases of SAT Solvers: ``On the Shoulders of Giants''}
\label{sec:codebases}

Progress in SAT solvers is often based on successful modifications of existing code-bases. 
One well-known tree of development is rooted in the code-base of \solver{Minisat} by E\'en and S\"orensson~\cite{Niklas:2003:Minisat}. 
A well-known fork of \solver{Minisat} is \solver{Glucose} by Audemard and Simon~\cite{Audemard:2018:Glucose} which introduced the influential literal block distance (LBD) for weighing clauses in forgetting heuristics~\cite{Audemard:2009:Glucose}. 
The SAT solver \solver{RISS} by Manthey is a successful and award-winning fork of \solver{Glucose} in combination with the \solver{Coprocessor}~\cite{Manthey:2012:Coprocessor2} preprocessor.

Another, more recent line of development in SAT solvers
is rooted in the \solver{CoMinisatPS} fork of \solver{Minisat} by Oh which introduced three-tier clause-management~\cite{Oh:2015:satunsat}. 
The SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heuristics at SAT Competition 2016~\cite{Liang:2016:LRB}. 
The then award-winning variant \solver{MapleCOMSPS} by Liang et al.~is an implementation of a hybrid branching heuristic of classic variable-state independent decaying sum (VSIDS)~\cite{Zhang:2001:ClauseLearning} and the newer learning rate based branching (LRB) in \solver{CoMinisatPS}~\cite{Liang:2016:MapleCOMSPS}.

In SAT Competition 2017, Luo et al.~integrated learned clause minimization based on unit propagation (\solver{LCM}) in their award-winning \solver{Maple\_LCM\_Dist}~\cite{Luo:2017:LCM} which also uses the new branching heuristic \solver{Distance (Dist)} in an initial solving period~\cite{Xiao:2017:MapleLCMDist}. 
Using \solver{Maple\_LCM\_Dist}, Ryvchin and Nadel successfully integrated conditional chronological backtracking (\solver{ChronoBT})~\cite{Nadel:2018:CBT}. 
They introduced their award-winning \solver{Maple\_LCM\_Dist\_ChronoBT} in SAT Competition~2018~\cite{Ryvchin:SC2018:MapleChronoBT}.

@@

Many independent and award-winning code-bases can be found among the SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award-winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} % \cite{Biere:SC2020} - or cite some older proceedings to match the "As of SAT Competition~2017" part?
by Biere is another independent, award-winning and structured implementation of the state-of-the-art in SAT solving. 
Its improved re-implementation \solver{Kissat}~\cite{Biere:SC2020} won in several of this competitions sequential tracks by a large margin.

The sequential variant of the \emph{new code-base} \solver{ParaFrost} by Osama and Wijs was submitted for the first time in this competition and introduced several interesting new concepts~\cite{Osama:SC2020:Parafrost}.

@@

Sequential SAT solvers have been evaluated in the Main, Planning and Incremental Library track of SAT Competition~2020. 
$18$ teams submitted a total of $48$ solvers and configurations to the Main track and the Planning track of the competition, and $4$ solvers participated the Incremental Library track. 
%and were ranked by the number of application categories in which they won based on their PAR-2 score. 
%Submissions to the Incremental Library track have been evaluated with respect to $6$ different applications -- each with $50$ benchmark instances. 
%
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
In the following sections, we outline the technical reports of the award winning solvers.

@@

Three configurations of \solver{Kissat} were evaluated in this competition, including one default configuration and two specialized configurations which are specifically tailored towards satisfiable and unsatisfiable instances, respectively. 
\solver{Kissat} received \emph{four awards}, by achieving the first place in the Main track, 
the best score on unsatisfiable instances, the second best score on satisfiable instances and the third place in the Planning track.

\solver{Kissat} is a low-level re-implementation of \solver{CaDiCaL} with new sophisticated lazy data-structures for clause state monitoring, e.g., through binary clause inlining, sentinel values and bit stuffing~\cite{Biere:SC2019,Biere:SC2020}. 
Moreover, forward subsumption for learned clauses is mostly replaced by vivification algorithms~\cite{ChuMinLi:2020:Vivification}. 
Since conflict-number has been observed to be too unstable for measuring the length of two alternating restart-modes, \solver{Kissat} uses the new unit ``ticks'' which approximates the number of cache-line accesses in unit-propagation~\cite{Biere:SC2020}.

\solver{Kissat} also exploits autarkies to account for saved phases. 
In order to keep valuable information of saved phases, before each rephasing step, \solver{Kissat} computes the largest autarky for the assignment implied by the current saved phases~\cite{Kiesl:2019:Autarkies}. 
As such an autarky might contain satisfying assignments which imply disconnected components, those variables are subject to subsequent variable elimination.

@@

\solver{CryptoMiniSat} received \emph{four awards}: the first place in the Incremental Library track,
the second place in the Planning track, the third place in the Main track, and the third-best score on satisfiable instances. 
Two submitted variants, \solver{default} and \solver{LSIDS}, scored mostly adjacent ranks in the evaluation of the tracks. 
The \solver{LSIDS} variant of \solver{CryptoMiniSat} comes with a new hybrid phase selection approach~\cite{Shaw:2020:LSIDS,Soos:SC2020}.

\solver{CryptoMiniSat} comes with an independent implementation of state-of-the-art hybrid branching heuristics which alternate between classic phase saving and target phase selection~\cite{Biere:SC2019}.
\solver{CryptoMiniSat-CCAnr} regularly schedules short periods of local search and imports the best assignment for phase selection---a procedure which is known as ``rephasing'' from CaDiCaL~\cite{Biere:SC2019}.
In addition, \solver{CryptoMiniSat-CCAnr} bumps the VSIDS scores of the first 100 variables in those clauses which the SLS solver weighs most hard to satisfy~\cite{Soos:SC2020}.

Inprocessing has been extended to include ternary resolution and more vivification~\cite{ChuMinLi:2020:Vivification}. 
\solver{CryptoMinisat} alternates decay factors of its branching heuristics, thus avoiding the restriction to a ``single best'' configuration~\cite{Soos:SC2020}. 
The submitted version of \solver{CryptoMinisat} entails a new optimized implementation of Gauss-Jordan Elimination~\cite{Soos:2020:CNFXOR}. 
\solver{CryptoMinisat} periodically executes the \solver{BreakId} algorithm to  calculate symmetry breaking clauses~\cite{Devriendt:2016:BreakId}.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} 
and an improved clause-learning \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful and won  first place in the Planning track.

@@

The \solver{Relaxed} fork of \solver{MapleLCMDistCBT-DL} was first presented in SAT Race 2019~\cite{Xindi:SC2019}. 
In SAT Competition~2020, its variant \solver{newTech} showed a good performance especially on satisfiable instances. 
The solver received two awards: second place in the Main track and the best score on satisfiable instances.

>>>>

In this section we will provide an overview of the participating teams and solvers, 
and summarize new strategies implemented in the best-performing solvers of the 2020 competition. 
We start with a few remarks on the evolution of code-bases of well-known SAT solvers. % in Section~\ref{sec:codebases}. 
%Our review of the winning strategies in this competition starts with sequential solvers in the Main, Planning, and Incremental Library tracks in Section~\ref{sec:part:seq},
%then we continue with winners of the Parallel track in Section~\ref{sec:part:par} and conclude with winning solvers in the Cloud track in Section~\ref{sec:part:cloud}.

\subsection{Evolution of SAT Solver Code-Bases: ``On the Shoulders of Giants''}
\label{sec:codebases}

Progress in SAT solvers is often based on successful modifications of existing and openly available solver code-bases. 
One well-known tree of code-base evaluation is rooted in the code-base of \solver{Minisat} by E\'en and S\"orensson~\cite{Niklas:2003:Minisat}. 
A well-known fork of \solver{Minisat} is \solver{Glucose} by Audemard and Simon~\cite{Audemard:2018:Glucose}. 
In particular, \solver{Glucose} introduced the influential literal block distance (LBD) heuristics for deciding which learned clauses to keep and which 
ones to forget during search~\cite{Audemard:2009:Glucose}. 
The SAT solver \solver{RISS} by Manthey is a further fork of \solver{Glucose}, combining \solver{Glucose}  with 
 the \solver{Coprocessor}~\cite{Manthey:2012:Coprocessor2} preprocessor.

A further, more recent line of evolution in SAT solvers
is rooted in the \solver{CoMinisatPS} by Oh, which is itself a again a fork of \solver{Minisat}, introduced three-tier clause-management~\cite{Oh:2015:satunsat}. 
Building on \solver{CoMinisatPS}, 
The SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heuristics at SAT Competition 2016~\cite{Liang:2016:LRB}. 
The at-the-time award-winning variant \solver{MapleCOMSPS} by Liang et al.~implements a hybrid branching heuristic of classic variable-state independent decaying sum (VSIDS)~\cite{Zhang:2001:ClauseLearning} and the newer learning rate based branching (LRB) of \solver{CoMinisatPS}~\cite{Liang:2016:MapleCOMSPS}.

For SAT Competition 2017, Luo et al.~integrated learned clause minimization based on unit propagation (\solver{LCM}) in their award-winning \solver{Maple\_LCM\_Dist}~\cite{Luo:2017:LCM} which also uses the new branching heuristic \solver{Distance (Dist)} in an initial solving period~\cite{Xiao:2017:MapleLCMDist}. 
Using \solver{Maple\_LCM\_Dist}, Ryvchin and Nadel successfully integrated conditional chronological backtracking (\solver{ChronoBT})~\cite{Nadel:2018:CBT}, with
the introduction of the award-winning solver \solver{Maple\_LCM\_Dist\_ChronoBT} in SAT Competition~2018~\cite{Ryvchin:SC2018:MapleChronoBT}.

@@

Many independent and award-winning code-bases can be found among the SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award-winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} % \cite{Biere:SC2020} - or cite some older proceedings to match the "As of SAT Competition~2017" part?
by Biere is another independent representative of state-of-the-art SAT solvers.
Its improved re-implementation \solver{Kissat}~\cite{Biere:SC2020} was quite successful in the 2020 competition.

The sequential variant of the \emph{new code-base} \solver{ParaFrost} by Osama and Wijs was submitted for the first time to the 2020 SAT Competition,
 introducing several interesting new concepts~\cite{Osama:SC2020:Parafrost}.

@@

Sequential SAT solvers have been evaluated in the Main, Planning and Incremental Library track of SAT Competition~2020. 
$18$ teams submitted a total of $48$ solvers and configurations to the Main track and the Planning track of the competition, and $4$ solvers participated the Incremental Library track. 
%and were ranked by the number of application categories in which they won based on their PAR-2 score. 
%Submissions to the Incremental Library track have been evaluated with respect to $6$ different applications -- each with $50$ benchmark instances. 
%
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
In the following, we provide a short overview of the best-performing solvers of 2020, based mainly on the solver descriptions submitted
to the 2020 competition proceedings by the authors of the individual solvers.

@@

Three configurations of \solver{Kissat} were submitted to the 2020  competition, including one default configuration and two specialized configurations which are specifically tailored towards satisfiable and unsatisfiable instances, respectively. 
\solver{Kissat} received \emph{four awards},achieving the first place in the Main track, 
the best score on unsatisfiable instances, the second-best score on satisfiable instances and the third place in the Planning track.

\solver{Kissat} is a low-level re-implementation of \solver{CaDiCaL} with new sophisticated lazy data-structures for clause state monitoring, e.g., through binary clause inlining, sentinel values and bit stuffing~\cite{Biere:SC2019,Biere:SC2020}. 
Moreover, forward subsumption for learned clauses is mostly replaced by vivification algorithms~\cite{ChuMinLi:2020:Vivification}. 
Since conflict number has been observed to be too unstable for measuring the length of two alternating restart modes, \solver{Kissat} uses the new unit ``ticks'' which approximates the number of cache-line accesses in unit-propagation~\cite{Biere:SC2020}. 
\solver{Kissat} also exploits autarkies to account for saved phases. 
In order to keep valuable information of saved phases, before each rephasing step \solver{Kissat} computes the largest autarky for the assignment implied by the current saved phases~\cite{Kiesl:2019:Autarkies}. 
As such an autarky might contain satisfying assignments which imply disconnected components, those variables are subject to subsequent variable elimination.

@@

\solver{CryptoMiniSat} received \emph{four awards}, achieving the first place in the Incremental Library track,
the second place in the Planning track, the third place in the Main track, and the third-best score on satisfiable instances. 
Two submitted variants, \solver{default} and \solver{LSIDS}, scored mostly adjacent ranks in the individual competition tracks.

The \solver{LSIDS} variant of \solver{CryptoMiniSat} comes with a new hybrid phase selection approach~\cite{Shaw:2020:LSIDS,Soos:SC2020}.
\solver{CryptoMiniSat} comes with an independent implementation of state-of-the-art hybrid branching heuristics which alternate between classic phase saving and target phase selection~\cite{Biere:SC2019}.
\solver{CryptoMiniSat-CCAnr} regularly schedules short periods of local search and imports the best assignment for phase selection---a procedure which is known as ``rephasing'' from CaDiCaL~\cite{Biere:SC2019}.
In addition, \solver{CryptoMiniSat-CCAnr} bumps the VSIDS scores of the first 100 variables in those clauses which the SLS solver weighs most hard to satisfy~\cite{Soos:SC2020}.
Inprocessing has been extended to include ternary resolution and more vivification~\cite{ChuMinLi:2020:Vivification}. 
\solver{CryptoMinisat} alternates decay factors of its branching heuristics, thus avoiding the restriction to a ``single best'' configuration~\cite{Soos:SC2020}. 
The submitted version of \solver{CryptoMinisat} entails a new optimized implementation of Gauss-Jordan Elimination~\cite{Soos:2020:CNFXOR}. 
\solver{CryptoMinisat} periodically executes the \solver{BreakId} algorithm to  calculate symmetry breaking clauses~\cite{Devriendt:2016:BreakId}.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} 
and an improved clause-learning \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful in the 2020 competition, achieving in particular the  first place in the Planning track.

@@

The \solver{Relaxed} fork of \solver{MapleLCMDistCBT-DL} was first presented in SAT Race 2019~\cite{Xindi:SC2019}. 
In SAT Competition~2020, its variant \solver{newTech} showed a good performance especially on satisfiable instances. 
The solver received two awards, achieving the second place in the Main track and the best score on satisfiable instances.
